\chapter{Introduction}

\section{Background}

The talk by Milanfar \cite{siam_slides_2016}, working at Google Research, about using the Graph Laplacian Operator for Image Processing purposes awakes curiosity.

Indeed, Milanfar reports that these techniques to build filters are used on smartphones, which means it is done in a reasonable time with limited computational resources.
Over 2 billion photos are shared daily on social media \cite{siam_slides_2016}, with very high resolutions and filters are often applied on them.
The algorithm must be excellent to be deployed at such scale.

We want to take a look into these spectral methods, see if they reveal systems of linear equations and what can be done with those dense pixel matrices with a fun application like image processing.

\section{Objective}

The objective of this thesis is to, in a first place, understand and implement the image processing algorithm shown by Milanfar's work on 2D images at Google.
We want to observe the performance and results of these algorithms to validate the usage of intrinsic spectral properties of an image.
We know that the spectrum of the graph associated to an image can serve for image segmentation.
Applied in a fine-grained manner, this segmentation can lead to image compression.
More classic algorithm can also be achieved with the described approach, such as image denoising, deblurring, sharpening, etc.

The next step includes extending these techniques on high-resolution and/or 3D images.
Given the results, new scalable processing approaches for 3D images could be explored.
These also include standard processing filters such as denoising and sharpening, but also segmentation and compression techniques.

These approaches, even though faster than other filters, still require heavy computations which can be numerous in the 3D case.
That is why we want to apply the algorithm on high performance architectures in order to achieve reasonable computation times.
The HPC component will help solving systems of linear equations with domain decomposition methods, which are naturally parallel methods, but also making the overall computations parallel.

\section{Related work}

\input{tex/1_3_related_work}

\section{Delimitations}

The number of investigated algorithms and variants depend on the pre-study and objectives that make the most sense for 3D images.
It is obviously also limited by the amount of time given to the degree project.

We will only consider image denoising in a first time, since it is the most straightforward application and the algorithm can be derived to many other applications.
